\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-09-20 gwnnDraft02.tex
First draft of conclusion

DylanLinthorne@cmai.carleton.ca \\

@@

The confinement of these sectors directly leads to the production of both baryons and mesons as we have the spontaneous breaking of SU$(6)_L \,\times$ SU$(6)_R \rightarrow\,$ SU$(6)$ and thus 35 pseudo-Goldstone bosons (pions). The masses obtained through the phase transition can be approximated through the use of a generalization of the Gell-Mann--Oakes--Renner relation~\cite{Gell-Mann,Schwartz:2013pla},

@@

A unpolarized stocastic gravitational wave background could be detectable if the the signal-to-noise ratio (SNR) is above some threshold value, $\rho > \rho_{th}$, dictated by the capabilities of future interferometers and pulsar timing arrarys (PTA). These interferometers/PTAs quote their experimental sensistivies in terms of spectral noise curves, $S_{\textrm{eff}}(f)$, which can be translated into units of energy density through $h^2 \Omega_{\textrm{eff}}(f) = \frac{2\pi^2}{3H^2}f^3 S_{\textrm{eff}}(f)$. If the experiment uses a single (multiple) detector, the auto-correlated (cross-correlated) SNR is used in comparing to the threshold value $\rho_{th}$. The auto-correlated and cross-correlated SNR are explictely given as~\cite{PhysRevD.59.102001},
\begin{equation}\label{eqn::SNR}
\begin{split}
\rho^2 =  \mathcal{T} \int_{f_{\textrm{min}}}^{f_{\textrm{max}}}\textrm{d}f \bigg( \frac{h^2 \Omega_{\textrm{GW}}(f)}{h^2 \Omega_{\textrm{eff}}(f)} \bigg)^2 \;\;\;\;\;\; \textrm{(auto-correlated)}
\\
\rho^2 = 2 \mathcal{T} \int_{f_{\textrm{min}}}^{f_{\textrm{max}}}\textrm{d}f \bigg( \frac{h^2 \Omega_{\textrm{GW}}(f)}{h^2 \Omega_{\textrm{eff}}(f)} \bigg)^2 \;\;\;\: \textrm{(cross-correlated)}
\end{split}
\end{equation}
Where $\mathcal{T}$ is the exposure time of the experiment. The integration covers the entire broadband range of frequencies $(f_{\textrm{min}}, f_{\textrm{max}})$. The Big Bang
Observer (BBO)~\citep{PhysRevD.72.083005} and the DECi-hertz Interferometer Gravitational wave Observatory (DEICIGO)~\citep{Sato_2017} would be built from an array of multiple interferometers, where as  the Laser Interferometer Space Antenna (LISA)~\citep{Audley:2017drz} and B-DEICIGO (scaled down DECIGO)~\citep{10.1093/ptep/pty078} are proposed to be single detector interferometers.

@@

The method of calculating the PLI curves involves plotting $h^2 \Omega_{\textrm{GW}}(f) $, using Eq.~(\ref{eqn:RelaDOF}), for various spectral indices $\beta \in (-8,-7,-6, ... , 6, 7, 8)$ and for some fixed threshold value of $\rho_{th}$. Curves generated in this way will lay tangent to the PSI curve, more formally,
\begin{equation}
h^2 \Omega_{\textrm{PLI}} = \max\limits_{\beta}\bigg[ h^2 \Omega_{\beta}\bigg(\frac{f}{f_{\textrm{ref}}}\bigg)^{\beta}  \bigg].
\end{equation}

The spectral noise curves used to create the PSI curves shown in figures~(\ref{fig::Nnatural}) and ~(\ref{fig:Haa}) where taken from~\citep{Robson_2019,PhysRevD.83.084036,doi:10.1142/S0218271813410137,10.1093/ptep/pty078, Breitbach:2018ddu} for the interferometers, and ~\citep{Breitbach:2018ddu, Janssen:2014dka} for the SKA pulsar timing array. We have assumed an observation time of $\mathcal{T} = 4 $ years for the interferometers and $\mathcal{T}= 5, 10, 20 $  years for the various stages of SKA.

@@

Gravitational wave signals continue to grow in importance as phenomenological si

We can think of these results providing 
This demonstrates the power of GW signals to probe deep into the unknown arena of complex hidden sectors --- whether there are 1, 5, or more hidden sectors

Further, should a stochastic GW background be detected, deviations from a standard power law curve would indicate a more complicated hidden sector.

Questions, of course, still remain: how can the estimations of $\alpha$ and $\beta \ H$ be improved?

>>>>

Dylan.linthorne@carleton.ca \\

@@

The confinement of these sectors directly leads to the production of both baryons and mesons as we have the spontaneous breaking of SU$(6) \,\times$ SU$(6) \rightarrow\,$ SU$(6)$ and thus 35 pseudo-Goldstone bosons (pions). The masses obtained through the phase transition can be approximated through the use of a generalization of the Gell-Mann--Oakes--Renner relation~\cite{Gell-Mann,Schwartz:2013pla},

@@

A unpolarized stocastic gravitational wave background could be detectable if the the signal-to-noise ratio (SNR) is above some threshold value, $\rho > \rho_{th}$, dictated by the capabilities of future interferometers and pulsar timing arrarys (PTA). These interferometers/PTAs quote their experimental sensistivies in terms of spectral noise curves, $S_{\textrm{eff}}(f)$, which can be translated into units of energy density through $h^2 \Omega_{\textrm{eff}}(f) = \frac{2\pi^2}{3H^2}f^3 S_{\textrm{eff}}(f)$. If the experiment uses a single (multiple) detector, the auto-correlated (cross-correlated) SNR is used in comparing to the threshold value $\rho_{th}$. The auto-correlated and cross-correlated SNR are explictely given as~\cite{PhysRevD.59.102001},
\begin{equation}\label{eqn::SNR}
\begin{split}
\rho^2 =  T \int_{f_{\textrm{min}}}^{f_{\textrm{max}}}\textrm{d}f \bigg( \frac{h^2 \Omega_{\textrm{GW}}(f)}{h^2 \Omega_{\textrm{eff}}(f)} \bigg)^2 \;\;\;\;\;\; \textrm{(auto-correlated)}
\\
\rho^2 = 2 T \int_{f_{\textrm{min}}}^{f_{\textrm{max}}}\textrm{d}f \bigg( \frac{h^2 \Omega_{\textrm{GW}}(f)}{h^2 \Omega_{\textrm{eff}}(f)} \bigg)^2 \;\;\;\: \textrm{(cross-correlated)}
\end{split}
\end{equation}
Where $T$ is the exposure time of the experiment. The integration covers the entire broadband range of frequencies $(f_{\textrm{min}}, f_{\textrm{max}})$. LISA~\citep{Audley:2017drz} and B-DEICIGO~\citep{10.1093/ptep/pty078} are proposed to be single detector interferometers, where as BBO~\citep{PhysRevD.72.083005} and DEICIGO~\citep{Sato_2017} would be built from an array of multiple interferometers.

@@

The method of calculating the PLI curves involves plotting $h^2 \Omega_{\textrm{GW}}(f) $, using Eq.~(\ref{eqn:RelaDOF}), for various spectral indices $\beta \in (-8,-7,-6, ... , 6, 7, 8)$ and for some fixed threshold value of $\rho_{th}$. Each curve will lay tangent to the PSI curve, more formally,
\begin{equation}
h^2 \Omega_{\textrm{PLI}} = \max\limits_{\beta}\bigg[ h^2 \Omega_{\beta}\bigg(\frac{f}{f_{\textrm{ref}}}\bigg)^{\beta}  \bigg].
\end{equation}

The spectral noise curves used to create the PSI curves shown in figures~(\ref{fig::Nnatural}) and ~(\ref{fig:Haa}) where taken from~\citep{Robson_2019,PhysRevD.83.084036,doi:10.1142/S0218271813410137,10.1093/ptep/pty078, Breitbach:2018ddu} for the interferometers, and ~\citep{Breitbach:2018ddu, Janssen:2014dka} for the SKA pulsar timing array. We have assumed an observation time of $T = 4 $ years for the interferometers and $T= 5, 10, 20 $  years for the various stages of SKA.

@@

As detection capabilities increase, gravitational wave signals continue to grow in importance as phenomenological signatures that can offer us a unique glimpse into the universe as it was in the early epochs. The space-based interferometers planned for the next generation of GW experiments will be sensitive enough to begin searching for signals of the cataclysmic disruption of space-time due to SFOPT. As we inch closer to these measurements becoming available, it becomes important to develop ways to analyze and understand this data.

Here, we examined several scenarios (including $N$naturalness) that involve multiple hidden sectors and calculated the GW profiles present. Our GW projections unambiguously demonstrate that scenarios with multiple hidden sector SFOPTs can begin to be probed by next generation experiments. Further, our results provide a framework for understanding and using GW signals in two opposite ways: first as a unique signal for specific theories featuring multiple SFOPTs and also as a way to explain signals that deviate from a standard power law spectrum.

In the former case, this demonstrates the power of GW signals to probe deep into the unknown arena of complex hidden sectors. Whether a model predicts of the presence of 1, 5, or more hidden sectors, any SFOPT can potentially contribute a GW signal and the sum of these signals can create a unique GW background; ultimately providing another means to test and constrain theories.

Shifting to the latter scenario, should a stochastic GW background be detected, deviations from a standard power law curve would indicate a more complicated hidden sector.  Explicitly, a broadening or distortion of the signal around the peak frequency could point to a multi-SFOPT scenario --- and gently guide us in the direction of multiple hidden sectors.

%Questions, of course, still abound: how can the %estimations of $\alpha$ and $\beta/H$ be improved?
